---
title: "GSS challange"
author: "kayleen hernandez!"
date: "October 5, 2021"
output: html_document
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(knitr)
GSS <-read.csv(file="GSScsv.csv",header=TRUE)
```

# Introduction

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?.  If you skimp in this section expect to be marked down in your grade. Paint a picture of the project a couple of paragraphs is expected!-->

Is attitude toward capitol punishment (CAPPUN) independent of race (RACE)?
If not, then whose view differs from whom??


$H_0$ : Ones attitude towards capital punishment is independent of race. 



$H_A$ : Ones attitude towards capital punishment is dependent of race.


# Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method. Discuss what data manipulations may be needed for example subsetting or other binning procedures and what sort of graphical displays and inferential tools will be used!-->


The data in this study is attitude towards capital punishment and the race of the individual. Both these variables used in this study are both considered categorical. 


# Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

## Descriptive Results

### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question. Bottom line is to explain the features of the graph that are relevant to conforming or departing to what is expected by your null hypothesis !-->



```{r}
dd2 <- GSS %>% group_by(RACE,CAPPUN) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))

basicC <- ggplot(dd2,aes(x=RACE,y=count,fill=CAPPUN))
basicC + geom_bar(stat="identity",position="dodge")
#Now for percentage plot

basicCC <- ggplot(dd2,aes(x=RACE,y=prcnt*100,fill=CAPPUN)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```

This graph shows that white and other races favor capitol punishment while blacks oppose.


### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
table1 <- xtabs(~RACE + CAPPUN, data=GSS)
rowPerc(table1)
colPerc(table1)
```

35% of blacks favor capital punishment. More than 65% of whites and other races favor capital punishment. More than 65% of blacks oppose capitol punishment It appears that there is a correlation between capital punishment and race.

## Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->

```{r}
fisherXY = function(x, y, confL = 0, printAll=FALSE){
  # Inputs: 
  #   x: factor variable, 2 levels
  #   y: factor variable, n levels
  # confL: confidence level, default = 0.95
  #        print only those tables, results if p-value < (1-confL)
  # printAll: overide the 'print only if p-value < (1-confL), and
  #           print all tables and fisher test results
  #
  # Constraint:
  #   x and y must be same length
  #
  # Outputs:
  #   All 2x2 Fisher exact tests
  
  if(confL<0 || confL> 1){
    print('error: confL must be between 0 and 1')
    return(999)
  }
  
  if(!is.factor(x)) x=as.factor(x)
  if(!is.factor(y)) y=as.factor(y)
  
  fisherTable = as.matrix(table(x, y))
  cat('\ncontingency table:\n')
  print(fisherTable)
  cat('\n***All 2X2 tests****\n\n')

  # all 2x2 tests
  levelsY = levels(y)
  levelsX = levels(x)
  
  m = length(levelsX)
  n = length(levelsY)
  
  confLevel = 0.95
  if (confL !=0 ) confLevel=confL
  
  for (i in 1:(m-1))
       for(j in ((i+1):m))
          for (k in 1:(n-1))
             for (l in ((k+1):n)) {
                
                 fisherIndexX = c(i,j)
                 fisherIndexY = c(k,l)
                 
                 fTempTable = as.matrix(fisherTable[fisherIndexX,
                                        fisherIndexY])
                 
                 fisherTest=fisher.test(fTempTable, conf.int = TRUE, 
                              conf.level = confLevel)
                 
                 if((fisherTest$p.value < (1-confLevel) ) || printAll){
                   
                       print(c(levelsX[i], levelsX[j]))
                       print(c(levelsY[k], levelsY[l]))
                       print(fTempTable)
                       print(fisherTest)
                       cat('\n')
                 }
                 
        }
}

```
```{r}
fisherXY(GSS$RACE,GSS$CAPPUN,printAll = T)
```

This data shows that between white and other races there is not much of a difference with a P-value of 0.27. There is an extreme difference between white and blacks with the P value of 0.0016.There is also an extreme difference between blacks and other races with the P value of 0.002022



# Discussion and Conclusion

The data shows that there is an extreme difference with the attitude toward capital punishment and races, However the difference is not evident with all races. Whites and other races are close in percentage of being in favor of capitol punishment.Moreover Blacks attitude toward capital punishment is a less percentage then both. In conclusion since the numbers do not correlate We reject the Null Hypothesis. 



<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variavbles causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->
